\papertitle

### 2014-03-17 final-report/paper.tex
Focus and context reference
Added fragment view

We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS). Our tool provides insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization. Our tool, which we call \system, helps developers understand and improve query execution, and provides insight into common problems such as data skew or performance bottlenecks.

@@

There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization. First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy. Once the logs are on one machine, the logs can be parsed and the data can be transformed into the desired format.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/viz_arch}
  \caption{Overview of the log collection and transformation architecture. Raw event logs are collected on each worker. The data is stored in relations. To download the data, a query has to be executed.}
  \label{fig:arch}
\end{figure}

@@

\subsection{Front-end}

The Myria web front-end server is written in Python and runs on Google App Engine. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3 to delive an interactive visualization experience to the user. D3 is a JavaScript framework for data visualziation on the web. We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.

@@

The \fragment view comprises two charts: the aggregation chart and the workers chart.

@@

A focus of the current implementation is on scalability to many workers and long running queries. More aggregations have to be done by the database itself. Most importantly, in the fragment visualization that shows spans for all events we should limit the data that is being downloaded and rendered to what the user can see. With the architecture as explained in Section~\ref{sec:back}, we are able to express these filter conditions as queries.

Besides performance improvements, we plan to integrate X-trace and its visualization to offer the orthogonal view and allow users to trace how tuple batches flow through the operators and between operators. Combined with the visualizations described in this paper, it will form a powerful debugging tool that will help us to better understand and improve query execution.

>>>>

We propose a visualization system for understanding and exploring query execution and data movement in a distributed database management system (DDBMS). Our tool will provide insight into: (1) data flow between query fragments and between workers, (2) query execution and operator dependencies, (3) cluster utilization, (4) network utilization. Our tool, which we call \system, will help developers understand and improve query execution, and will provide insight into common problems such as data skew or performance bottlenecks.

@@

There are two possible approaches that we explored to collect event logs that can be used to extract spans and other performance related data necessary for the visualization. First, events could be logged to files using a standard logging tool and then collect the files on one node using remote copy. Once the logs are in one place, they can be parsed and the data transformed into the desired format.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/viz_arch}
  \caption{Architecture}
  \label{fig:arch}
\end{figure}

@@

\subsection{Front-end}
The Myria web front-end server is written in Python and runs on Google App Engine. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3 to delive an interactive visualization experience to the user. D3 is a JavaScript framework for data visualziation on the web. We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.

@@

The \fragment view comprises two charts: the utilization chart and the
operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows the utilization
of the cluster over time, inside the selected fragment. The chart is used to
quickly reveal any patterns in the schedule, like tail latency, long periods of
idleness, etc. We implemented a small brush (at the top of the figure) that
allows the user to zoom-in and further analyze the problematic area.

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/utilization_chart}
  \caption{The utilization chart. }
  \label{fig:utilization_chart}
\end{figure}

The utilization chart also implements a brush that allows the user to further
analyze the query execution by displaying the schedule of each of the worker
nodes, in the operators chart (Fig.~\ref{fig:operators_chart}). The per-worker
schedule shows what operators the worker executed and for how long.

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/operators_chart}
  \caption{The operators chart.}
  \label{fig:operators_chart}
\end{figure}

@@

% Dom
